Optimal recommender system with strategic alternatives \\ Market Design. Term Project

### 2020-01-14 Draft.tex
draft

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=15mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{definition}[theorem]{Definition}
\title{Optimal recommender system with strategic alternatives \\ Market Design. Term Project}
\date{}
\author{Danil Fedchenko}
\begin{document} % 
	\maketitle
	\section{Introduction}
	Suppose you would like to find a restaurant to go out tonight what would you do? 
	Recommender systems became an integral part of our lives. The most widely known examples of such a kind of services include: \textit{TripAdvisor} which presents ranking on Hotels and Restaurants for customers based on previous users' experience; \textit{Yelp} which provides a forum with reviews and ranking for different businesses; \textit{Netflix} which recommends movies and series also based on information aggregated from different customers. All those services are very important since, on one hand, they provide people with information about options they otherwise may never be aware of, at no (or very low) cost. On the other hand, these services aggregate the information and may make rather precise recommendations to people which significantly reduce search cost for them. But here the interesting question arises: how to disclose the information, the platform aggregated, optimally? It turns out that the full disclosure may not be optimal from the social welfare point of view. This happens since users, who do not internalize the information externalities they impose to other agents, underexplore i.e. start to choose the safe option and do not try new things.

	Several papers studied this question in different frameworks (see Literature review section) and the main conclusion that all those paper draw is that there should be an optimal level of information obfuscation in order to induce short-living and myopic users to experiment. However, in the previous works the quality of alternatives was supposed to be exogenous to the process of recommendation and learning. So, the recommender policy the platform implements cannot affect the realized quality of alternatives that users choose. But in reality, these alternatives are also strategic agents (think about restaurants or hotels), and in most cases these agents are way more sophisticated than users. These strategic agents may adjust the quality of the good to the recommender policy the platform is employing. And the platform, in turn, should take the strategic nature of alternatives into account once developing the policy. To my best knowledge, there are no papers that study the optimal recommender policy with strategic alternative, so this work may be considered as the first small step in this direction.

	I build a model that embraces three main parties: the platform, the firm and users. Users are choosing between the good of \textit{ex-ante} unknown quality provided by the firm and a safe outside option. The firm can exert costly private effort to increase the quality of the service and the platform designs the optimal recommender policy taking this into account. I demonstrate that the optimal policy is different once the firm becomes strategic. Also, I derive the optimal recommender policy in the class of simple "1-round" policies. The rest of the paper is organized as follows: the next section presents the literature review, section 3 describes the model, section 4 contains the main results, and last three sections embrace a brief discussion, possible extensions and directions for the future research, and a short conclusion.
	\section{Literature Review}
	This work is related to the strand of literature on recommender systems. One of the first papers which studied the optimal design of recommender system is the work by Kremer et al \cite{kremer2014}. The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff. One of the alternatives is \textit{a priori} better, in the sense that the expected reward it generates is higher. The finite number of short-living consumers arrive sequentially and each chooses between either of two alternatives. The platform accumulates review and commits to the recommendation policy it uses to make recommendations to the newcomers. The papers, firstly, neatly in simple terms demonstrates why the full revelation policy may be suboptimal. Secondly, the optimal recommender policy is derived which has a "threshold" nature. That is, \textit{a priori} better alternative is recommended if its quality turns out to be higher than the certain threshold. Otherwise the other alternative is recommended, and then once both alternatives were tried the best one is recommended until the end. The setup of this model is very similar to my, however, it has a clear distinction. In my model the quality of the alternative is endogenous to the process of recommendation. This suggests that the threshold-like type of policies cannot be optimal in the presence of strategic alternatives because once the firm knows that it will be recommended forever, it does not have any incentives to maintain the high quality of the service it provides.

	Another relevant paper is that of Che and Horner \cite{che2015}. In the paper a continuous flow of customers arrives, and each should choose between the good or an outside option. The quality of the good is exogenously predetermined but unknown. Customers who decide to consume the good send noisy signal about its quality to the platform. The signal is noisy in the sense that even if the quality of the good is good the information about it reaches to the platform only with some probability (different from 1). The platform updates its beliefs about the quality and tailors recommender policy contingent on the beliefs. Authors demonstrate that the full revelation policy is suboptimal and find the optimal policy which again has a threshold nature. Namely, once the platform receives the signal that the quality is good, it will recommend the good until the end. Otherwise, if no news arrive it is optimal to experiment with the risky alternative (recommend it) until the beliefs that its quality is good fall below the certain threshold. After that the optimal policy never recommends this alternative.

	Finally, Papanstasiou et al \cite{papanastasiou2017} extended ideas developed in \cite{kremer2014} and \cite{che2015} and build a dynamic, stochastic model. In their model, agents also arrive sequentially to the platform and should choose between two alternatives that generate stochastically exogenous payoff. The platform in the paper designs a recommendation policy. Although the optimal policy turns out to have a complicated form, authors demonstrate that in this case the full revelation policy is not the best one, and the optimal policy involves information obfuscation.

	The common feature of all aforementioned models is some form of exogeneity of the quality of the alternative. In the next section I present a model with the strategic firm that can adjust the quality of the good to the learning process and recommendation policy.

@@

\bibliography{lib}{}
\bibliographystyle{unsrt}
\end{document}

>>>>

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{breakcites}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=15mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{definition}[theorem]{Definition}
\title{Optimal recommender system with strategic alternatives \\ Market Design. Term Project}
\date{}
\author{Danil Fedchenko}
\begin{document} % 
	\maketitle
	\section{Introduction}
	Suppose you came to New-York and would like to find a good restaurant to go out tonight. Or you are going to spend your upcoming vacation on the coast of Spain and you need to find a hotel. What would you do? Where would you search for?  Before the World Wide Web, people in such cases relied upon guide-books and brochures distributed by traveling agents. Those guides usually quickly became out-of-dated and were criticized for containing a misleading information.

	After the advent of the Internet, finding up-to-date information has become much easier. To make this process even easier there began to appear Internet platforms that aggregate customers' reviews on products and service providers. Some of that platforms simply collect the reviews and present a rating that is just an average measure of consumers' satisfaction (weighted or simple mean of rates). Examples embrace \textit{Yelp} for restaurants, home and car services, or \textit{SiteJabber} for online businesses.

	 Other platforms came further and began to not only collect the reviews but also to make individual recommendations based on these reviews. Usually these recommendations have a form of ranking of several alternatives that are ranked according to the secret propriety algorithm that the particular platform employs. The most widely known examples of such a kind of platforms include: \textit{TripAdvisor} which presents ranking on Hotels and Restaurants based on previous users' experience; \textit{Trustpilot} which is one of the most popular review services for a large variety of businesses; \textit{Glassdoor} that uses reviews of employees to rank companies by workers' satisfaction; and many other examples. Also, the interesting example is \textit{Amazon} which before 2015 calculated ranking just as an average of customers' reviews but after 2015 it changed the algorithm to a more complex, secret algorithm. The reason why \textit{Amazon} may have done so will be highlighted later in the paper.

	 In this work by recommender system I mean such a kind of platforms that aggregates reviews of customers and makes a recommendations to newcomers based on these reviews. Those recommendation may be direct or indirect (e.g. the platform can provide a ranking of alternatives, this is not the direct recommendation to choose the top-ranked alternative, however, users understand that the platform thinks that this alternative is the best fit for them, so the platform implicitly recommends this alternative).

	Recommender systems became an integral part of our lives. Those services are very popular since, on one hand, they provide people with information about options they otherwise may never be aware of, at no (or very low) cost. On the other hand, these services aggregate the information and may make rather accurate recommendations to people which significantly reduces search cost for them. But here the interesting question arises: how to disclose the information, the platform aggregated, optimally? Is it the best option from the social welfare standpoint to disclose all the information?

	Several papers studied this question in different frameworks (see Literature review section) and using different models. It turns out that the full disclosure may not be optimal. This happens since users, who do not internalize the information externalities they impose to other agents, underexplore i.e. start to choose the safe option and do not try new things. The main conclusion that all those paper draw is that there should be an optimal level of information obfuscation in order to induce users to experiment. However, in the previous works the quality of alternatives was supposed to be exogenous to the process of recommendation and learning. So, the recommender policy the platform implements cannot affect the realized quality of alternatives that users choose. But in reality, these alternatives may be also strategic agents (think about restaurants, hotels or any other businesses), and in most cases these agents are way more sophisticated than users who usually come to the platform choose a restaurant or a hotel and then leave the platform for a long time. These strategic alternatives may adjust the quality of the good to the recommender policy the platform is employing. And the platform, in turn, should take the strategic nature of alternatives into account once developing its policy. The main question I am interested to address here is: what should be the best recommender policy the platform may implement if it deals with strategic alternatives which can adjust the quality of provided service to the recommender policy. To my best knowledge, there are no papers that study the optimal recommender policy in this framework, so the model I present here may be considered as the first small step in this direction.

	I build a model that embraces three main parties: the platform, the firm and users (customers). Users arrive each period and are choosing between the good which is supplied by the firm and the outside option. Customers differ in the value of the outside option, and the utility they derive upon choosing the outside option is known for each customer. At the same time, the value customers derive from consuming the good is \textit{ex-ante} unknown since the good can have different quality and customers are unaware of the quality before making a choice. The firm can exert costly private effort to increase the quality of the good and the platform designs the optimal recommender policy taking this into account. I demonstrate that the optimal policy is different once the firm becomes strategic. Also, I derive the optimal recommender policy in the class of simple "1-round" policies. Additionally, I analyze how this policy changes if the customers also are strategic and follow the recommendation the platform makes if and only if the expected utility they may derive exceeds the value of the outside option. The results presented here highlight a new perspective on the role of the recommender system: it can not only aggregate information and provides the consumer side with reliable reviews and recommendations but also to encourage the producer side to invest in the product quality and maintain the high level of it.

	The rest of the paper is organized as follows: the next section presents the literature review, section 3 describes the model, section 4 contains the main results, and last three sections embrace a brief discussion, possible extensions and directions for the future research, and a short conclusion.
	\section{Literature Review}
	This work is related to the strand of literature on recommender systems. The first prototype of the recommender system has been created by Xerox Palo Alto Research Center in the late 1980-s early 1990-s and was described by  \cite{goldberg1992using}. This system has been called \textit{Tapestry} and has been designed in order to filter out a skyrocketed amount of the incoming emails. The system was trying to make the incoming stream of emails interesting for the particular person based on whether these emails were interesting for people whose interests are close to the interests of this person. Since that time many papers has been written that mainly explore the recommender systems from the Computer Science standpoint (see \cite{Beel2015r} for the recent survey).

	One of the first papers which highlights the issue from the point of view of Economics and studies the optimal design of recommender system is the work by \cite{kremer2014}. The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff. One of the alternatives is \textit{a priori} better, in the sense that the expected reward it generates is higher. The finite number of short-living consumers arrive sequentially and each chooses between either of two alternatives. The platform accumulates review and commits to the recommendation policy it uses to make recommendations to the newcomers. The papers, firstly, neatly in simple terms demonstrates why the full revelation policy may be suboptimal\footnote{Maybe the appearance of this paper and similar to that in 2014 was the reason why \textit{Amazon} decided to switch their recommendation policy from the fully revealing one.}. Secondly, the optimal recommender policy is derived which has a "threshold" nature. That is, \textit{a priori} better alternative is recommended if its quality turns out to be higher than the certain threshold. Otherwise the other alternative is recommended, and then once both alternatives were tried the best one is recommended until the end. The model I build and present in the next section has similar setup, however I allow the quality of the alternative to be exogenous to the process of learning. This suggests that the threshold-like type of policies similar to those obtained in \cite{kremer2014} may not be optimal in the presence of strategic alternatives because once the firm knows that it will be recommended forever, it does not have any incentives to maintain the high quality of the service it provides.

	Another relevant paper is that of Che and Horner \cite{che2015}. In the paper a continuous flow of customers arrives, and each should choose between the good or an outside option. The quality of the good is exogenously predetermined but unknown. Customers who decide to consume the good send a noisy signal about its quality to the platform. The signal is noisy in the sense that even if the quality of the good is good the information about it reaches to the platform only with some probability (different from 1). The platform updates its beliefs about the quality and tailors a recommender policy contingent on the beliefs. Authors demonstrate that the full revelation policy is suboptimal and find the optimal policy which again has a threshold nature. Namely, once the platform receives the signal that the quality is good, it will recommend the good until the end. Otherwise, if no news arrive it is optimal to experiment with the risky alternative (recommend it) until the beliefs that its quality is good fall below the certain threshold. After that, the optimal policy prescribes to never recommend this alternative.

	Finally, Papanstasiou et al. \cite{papanastasiou2017} extended ideas developed in \cite{kremer2014} and \cite{che2015}, and build a dynamic, stochastic model. In their model, agents also arrive sequentially to the platform and should choose between two alternatives that generate stochastically exogenous payoff. The platform in the paper designs a recommendation policy. Although the optimal policy turns out to have a complicated form, authors demonstrate that in this case the full revelation policy is not the best one, and the optimal policy involves information obfuscation.

	The model that will be considered lately in the paper includes the stream of agents who should take an action and the results of their action may affect the actions of subsequently arriving agents. This feature may links this work to the information cascading literature (seminal papers are those of \cite{banerjee1992simple} and \cite{bikhchandani1992theory}). The information cascading in my model may happen if the platform reveals the information of the previous users' experience to the newcomers. However the scope of this work is completely different: the platform designs a recommender policy that prevents the information cascading from happening and moreover this policy induces the firm to exert effort and maintain a high quality of the good.

	The common feature of the models with recommender systems considered in the aforementioned papers is some form of exogeneity of the quality of the alternative: in \cite{kremer2014} and \cite{che2015} it is determined and fixed, while in \cite{papanastasiou2017} it realizes randomly according to the parametric family of distribution with unknown parameters. In the next section I present a model with the strategic firm that can adjust the quality of the good to the learning process and recommendation policy.

@@

\newpage
\bibliography{lib}{}
%\bibliographystyle{unsrt}
\bibliographystyle{apalike}
\end{document}
